Reduced plasticity and opportunity for selection in terrestrial ectotherm populations under climate change

### 2023-12-04 ms.qmd
Added new refs

#### Affliations:
$^1$ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2600, Australia
$^2$ Ecology and Evolution Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, NSW, Australia
$^3$ SOLES, University of Sydney, Sydney, NSW, Australia
$^4$ Department of Biology, Lancaster University, Liverpool, UK

@@

Climate change is expected to result in warmer and more variable thermal environments for many organisms globally [@Suarez2023; @Easterling2000; @Ummenhofer2017]. Greater thermal variability in the past is expected to have resulted in strong selection pressures leading to genetic adaptation and/or the evolution of adaptive phenotypic plasticity -- both of which are considered important for population resilience to contemporary human-induced climate change [@Seebacher2015-wi; @Merila2014; @Chevin2015; @Chevin2017; @Chevin2010; @Nunney2016]. The extent to which adaptive plasticity and genetic adaptation will allow organisms to adapt to new environmental conditions is not yet clearly established [@Chevin2017]. Without such responses, high rates of extinction are predicted if organisms cannot track suitable habitats [@Cahill2012; @Nunney2016]. Phenotypic plasticity is expected to the the 'first line of defence' against changing climates by buying time for genetic adaptation to take place (i.e., the 'plasticty first hypothesis' - ). Phenotypic plasticity is predicted to evolve when environmental variability is high but predictable and the costs of plasticity is low [@Chevin2017; @Scheiner2020; @Nunney2016]. Despite this theoretical expectation, empirical support is scant [], likely because many organisms can behaviorally adjust micro-habitat selection to offset thermal stress, the costs of plasticity are high [@Chevin2015], and/or the prediction is likely only supported for specific plastic responses (i.e., active and developmental plasticity). Reversible forms of phenotypic plasticity, such as acclimatization, may not be expected to adhere to such predictions.

>>>>

#### Affliations:
$^1$ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2600, Australia \n
$^2$ Ecology and Evolution Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, NSW, Australia \n
$^3$ SOLES, University of Sydney, Sydney, NSW, Australia \n
$^4$ Department of Biology, Lancaster University, Liverpool, UK

@@

Climate change is expected to result in warmer and more variable thermal environments for many organisms globally [@Suarez2023; @Easterling2000; @Ummenhofer2017]. Greater thermal variability in the past is expected to have resulted in strong selection pressures leading to genetic adaptation and/or the evolution of adaptive phenotypic plasticity -- both of which are considered important for population resilience to contemporary human-induced climate change [@Seebacher2015-wi; @Merila2014; @Chevin2015; @Chevin2017; @Chevin2010; @Nunney2016]. The extent to which adaptive plasticity and genetic adaptation will allow organisms to adapt to new environmental conditions is not yet clearly established [@Chevin2017]. Without such responses, high rates of extinction are predicted if organisms cannot track suitable habitats [@Cahill2012; @Nunney2016]. Phenotypic plasticity is expected to the the 'first line of defence' against changing climates by buying time for genetic adaptation to take place [i.e., the 'plasticty first hypothesis' - @WestEberhard2003, see also @Lande2009]. Phenotypic plasticity is predicted to evolve when environmental variability is high but predictable and the costs of plasticity is low [@Chevin2017; @Scheiner2020; @Nunney2016]. Despite this theoretical expectation, empirical support is scant [], likely because many organisms can behaviorally adjust micro-habitat selection to offset thermal stress, the costs of plasticity are high [@Chevin2015], and/or the prediction is likely only supported for specific plastic responses (i.e., active and developmental plasticity). Reversible forms of phenotypic plasticity, such as acclimatization, may not be expected to adhere to such predictions.
